---
title: "Lattice"
output: html_document
---
Lattice is kinda more for plotting many things at once. A lot of code is implemented in the
`grid` package. Lattice package does not adhere to semantics of plotting data first and then
overlaying it with styles. Everything is done in one funciton call.

Functions:

* `xyplot` scatterplots
* `bwplot` box and whiskers plot
* `histogram` duh
* `stripplot` like boxplot, but with actual points
* `dotplot` plots dots on violin-like strings
* `splom` scatterplot matrix
* `levelplot, contourplot` for plotting image data

Lattice functions generally take formula style

`xyplot(y~x|f*g, data)` left goes y, x goes right,
f and g are conditional variables, the plot literally means "I would like to look at y and x 
for any level of f and g". `data` contains the actual data for the plot.

```{r}
library(lattice)
library(datasets)
xyplot(Ozone~Wind, data=airquality)
```

```{r}
library(lattice)
library(datasets)
airquality<-transform(airquality, Month=factor(Month))
xyplot(Ozone~Wind|Month, data=airquality, layout=c(5,1))
```

Lattice functions return objects of type *trellis*. Needs a `print` call to be sent to a device. Most of the time you do not have to worry about the plotting device, R will take care of it.

```{r}
p<-xyplot(Ozone~Wind, data=airquality) #nothing happens
print(p)
```
Lattice function has a panel funciton that controls what's happening incide every panel of the plot.

```{r}
set.seed(10)
x<-rnorm(100)
f<-rep(0:1, ecah=50)
y=x+f-f*x+rnorm(100, sd=0.5)
f<-factor(f, labels=c("Group 1","Group 2"))
xyplot(y~x|f, layout=c(2,1)) # Plot with 2 panels
```

```{r}
xyplot(y~x|f, panel=function(x,y,...){
  panel.xyplot(x, y, ...) #first call the function default
  panel.abline(h=median(y),lty=2) #Add horyzontal line at the median
})
```

```{r}
xyplot(y~x|f, panel=function(x,y,...){
  panel.xyplot(x, y, ...) #first call the function default
  panel.lmline(x,y, col=2) #Overlay a simple linear regerssion
})
```
Can't use any base funcitons for annotation.

#ggplot2
Built on top of `grid` package.
`qplot` is an analog of the base plotting system. The main difference is that if you pass some data to the qplot, you have to specify where is the data coming from. If a dataframe with the data, is not specified, it will look over the environment.

**aestetics** are the things like size, colors etc.
**geoms** specify what type of plots are being used.

Factors are important to subset the datasets. Factors should be labeled.
`ggblot` is the core funciton of the package.

```{r}
library(ggplot2)
```
A Hello World plot:
```{r}
qplot(displ, hwy, data=mpg)
```
```{r}
qplot(displ, hwy, data=mpg, color=drv)
```
```{r}
qplot(displ, hwy, data=mpg, geom=c("point", "smooth"))
```
If only one variable is specified -> a histogram is generated
```{r}
qplot(hwy, data=mpg, fill=drv)
```
##Facets
```{r}
qplot(displ,hwy, data=mpg, facets=.~drv)
qplot(hwy, data=mpg, facets=drv~., binwidth=2)
```

Main components of the ggplot graphics

* dataframe
* aestetic mapping: how data are mapped to colours and shapes
* facets for conditional plots
* stats for binning and quantile smoothing
* scales: what scales and estetics mappings use
* coordinate system

Plots are built like layers:

- plot the data
- overlay summary
- metadata and annotation

Can always call a `summary(ggplot.object)` to see the components: data, mapping and faceting.

##Annotation

* labels `xlab()`, `ylab()`, `labs()` and `ggtitle()`
* each of the geom functions have parameters to modify
* for the things that only make sence globaly use `theme()`
* two standard appearences are included `theme_gray()` and `theme_bw()`

```{r}
expression("log"* PM[2.5])
```







