---
title: "Plotting Systems In R"
output: html_document
---

#Base
```{r}
library(datasets)
hist(airquality$Ozone)
with(airquality, plot(Wind, Ozone))
airquality<-transform(airquality,Month=factor(Month))
boxplot(Ozone~Month,airquality,lab="Month",ylab="Ozone (ppb)")
```

##Parameters

- `pch`: plotting system with default as open cicle
- `lty`: the line type with a default of solid line, but can be dashed, dotted, etc
- `lwd`: line width, specified as integer multiple
- `col`: color, specified as number, string, or hex code, `colors()` function gives color by name
- `xlab`: character vector for the x-axis
- `ylab`: character vector for the y-axis

- `par()` is a global function that affects all the plots in R base graphics
- `las`: orientation of the axis on the plot
- `bg`: background, duh!
- `mar`: the margin size
- `oma`: the outer margin size (default is 0 for all sides)
- `mfrow`: number of plots per row, column plots are filled row-wise
- `mfcol`: number of plots per column --//--

Can call the `par()` with different parameters in order to see current defaults
```{r}
par("lty")
par("col")
par("pch")
```

#Types of `{base}` plots
- `plot`: make a scatterplot or other plot, depending on the type of object passed
- `lines`: add lines
- `points`: add points
- `text`: add text
- `title`: add title
- `mtext`: add library text to the margins
- `axis`: add axis ticks/labels

```{R}
with(airquality, plot(Wind, Ozone, main= "Ozone blabla"))
with(subset(airquality,Month==5), points(Wind,Ozone, col="blue"))
with(subset(airquality,Month!=5), points(Wind,Ozone, col="red"))
legend("topright",pch=1,col=c("blue","red"),legend= c("May","Not May"))

with(airquality, plot(Wind, Ozone, main= "Ozone blabla",pch=20))
model<-lm(Ozone~Wind,airquality)
abline(model,lwd=2)

par(mfrow=c(1,3),mar=c(4,4,2,1), oma=c(0,0,2,0))
with(airquality, {
  plot(Wind, Ozone, main="Ozone and Wind")
  plot(Solar.R,Ozone,main="Ozone and Solar Radiation")
  plot(Temp,Ozone,main="Ozone and Temperature")
  mtext("Ozone and Weather blable", outer = TRUE)
})
```
