---
title: "PlottingSystemsInR"
output: html_document
---

#Base
```{r}
library(datasets)
hist(airquality$Ozone)
with(airquality, plot(Wind, Ozone))
airquality<-transform(airquality,Month=factor(Month))
boxplot(Ozone~Month,airquality,lab="Month",ylab="Ozone (ppb)")
```

##Parameters

- `pch`: plotting system with default as open cicle
- `lty`: the line type with a default of solid line, but can be dashed, dotted, etc
- `lwd`: line width, specified as integer multiple
- `col`: color, specified as number, string, or hex code, `colors()` function gives color by name
- `xlab`: character vector for the x-axis
- `ylab`: character vector for the y-axis

- `par()` is a global function that affects all the plots in R base graphics
- `las`: orientation of the axis on the plot
- `bg`: background, duh!
- `mar`: the margin size
- `oma`: the outer margin size (default is 0 for all sides)
- `mfrow`: number of plots per row, column plots are filled row-wise
- `mfcol`: number of plots per column --//--

Can call the `par()` with different parameters in order to see current defaults
```{r}
par("lty")
par("col")
par("pch")
```
