---
title: "Hierarchical clustering"
output: html_document
---

An agglomerative approach:

* find two closest things
* put them together
* find next closest

Requires:

* a defined distance
* agglomerating method

Produces:

* a dendrogram/tree

```{r}
set.seed(1234)
par(mar=c(0,0,0,0))
x=rnorm(12, mean=rep(1:3,each=4),sd=0.2)
y=rnorm(12, mean=rep(c(1,2,1),each=4),sd=0.2)
plot(x,y,col="blue",pch=19,cex=2)
text(x+0.05,y+0.05,labels=as.character(1:12))
dataFrame<-data.frame(x=x,y=y)
d<-dist(dataFrame)
hc<-hclust(d)
plot(hc)
```

##Heatmap

```{r}
dataFrame<-data.frame(x=x,y=y)
set.seed(143)
dataMatrix<-as.matrix(dataFrame)[sample(1:12),]
heatmap(dataMatrix)
```
Rafa's distance clustering

##K-means clustering

* fixed number of clusters
* gets centroids for each cluster
* assigns things to the closest centroid
* recalculates centroids (basically is a form of MLE)

Requires:

* a defined distance metric
* a number of clusters
* an intial guess as to cluster centroids

Produces:

* final cluster centroid estimation
* an assignment of each point to clusters

```{r}
km<-kmeans(dataFrame, centers = 3)
names(km)
km$cluster
plot(x, y, col=km$cluster, pch=19,cex=2)
points(km$centers, col=1:3,pch=3, cex=3,lwd=3)
par(mfrow=c(1,2), mar=c(2, 4, 0.1, 0.1))
image(t(dataMatrix)[,nrow(dataMatrix):1],yaxt="n")
image(t(dataMatrix)[,order(km$cluster)],yaxt="n")
```

Kmeans is not determenistic. May be a good idea to run the algorithm a number of times to see if it converges to the same centroids (approximately)

#Dimension Reduction
##Principal Components Analysis
```{r}
set.seed(678910)
dataMatrix<-matrix(rnorm(400),nrow=40)
for (i in 1:40){
  coinFlip<-rbinom(1, size=1,prob=0.5)
  if(coinFlip){
    dataMatrix[i,]<-dataMatrix[i,]+rep(c(0,3),each=5)
  }
}
par(mar=rep(0.2,4))
image(1:10,1:40,t(dataMatrix)[,nrow(dataMatrix):1])
heatmap(dataMatrix)
hh <- hclust(dist(dataMatrix))
dataMatrixOrdered <- dataMatrix[hh$order, ]
par(mfrow = c(1, 3))
image(t(dataMatrixOrdered)[, nrow(dataMatrixOrdered):1])
plot(rowMeans(dataMatrixOrdered), 40:1, , xlab = "Row Mean", ylab = "Row", pch = 19)
plot(colMeans(dataMatrixOrdered), xlab = "Column", ylab = "Column Mean", pch = 19)
```

##SVD

If X is a matrix with each variable in a column and each observation in a row then the SVD is a "matrix decomposition"

X=UDV^T^
where the columns of U are orthogonal (left singular vectors), the columns of V are orthogonal (right singular vectors) and D is a diagonal matrix (singular values).

##PCA

The principal components are equal to the right singular values if you first scale (subtract the mean, divide by the standard deviation) the variables.

```{r}
svd1 <- svd(scale(dataMatrixOrdered))
par(mfrow = c(1, 3))
image(t(dataMatrixOrdered)[, nrow(dataMatrixOrdered):1])
plot(svd1$u[, 1], 40:1, , xlab = "Row", ylab = "First left singular vector",
pch = 19)
plot(svd1$v[, 1], xlab = "Column", ylab = "First right singular vector", pch = 19)
```

###Variance explained
```{r}
par(mfrow = c(1, 2))
plot(svd1$d, xlab = "Column", ylab = "Singular value", pch = 19) 
plot(svd1$d^2/sum(svd1$d^2), xlab = "Column", ylab = "Prop. of variance explained",
pch = 19)
```

##Missing values
```{r}
library(impute) ## Available from http://bioconductor.org 
dataMatrix2 <- dataMatrixOrdered 
dataMatrix2[sample(1:100,size=40,replace=FALSE)] <- NA 
dataMatrix2 <- impute.knn(dataMatrix2)$data
svd1 <- svd(scale(dataMatrixOrdered))
svd2 <- svd(scale(dataMatrix2)) 
par(mfrow=c(1,2)) 
plot(svd1$v[,1],pch=19); plot(svd2$v[,1],pch=19)
```

#Colours

`library(grDevices)`

* The grDevices package has two functions 
-- `colorRamp`
-- `colorRampPalette`
* These functions take palettes of colors and help to interpolate between the colors
* The function `colors()` lists the names of colors you can use in any plotting function

##Colour Palette Utilities in R

* `colorRamp`: Take a palette of colors and return a function that takes valeus between 0 and 1, indicating the extremes of the color palette (e.g. see the 'gray' function)
* `colorRampPalette`: Take a palette of colors and return a function that takes integer arguments and returns a vector of colors interpolating the palette (like heat.colors or topo.colors)

```{r}
pal <- colorRamp(c("red", "blue"))
pal(0)
pal(1)
pal(0.5)
pal(seq(0, 1, len = 10))
pal <- colorRampPalette(c("red", "yellow"))
pal(2)
pal(10)
```
##`RColorBrewer Package`

One package on CRAN that contains interes1ng/useful color palettes

* There are 3 types of palettes
 - Sequential 
 - Diverging
 - Qualitative
* Palette informa1on can be used in conjunction with the `colorRamp()` and `colorRampPalette()`
```{r}
library(RColorBrewer)
cols <- brewer.pal(3, "BuGn")
cols
pal <- colorRampPalette(cols)
image(volcano, col = pal(20))
x<-rnorm(10000)
y<-rnorm(10000)
smoothScatter(x,y)
```

The `rgb` function can be used to produce any color via red, green, blue proportions

* Color transparency can be added via the alpha parameter to rgb
* The colorspace package can be used for a different control over colors







